---
title: |
    | Coursera Finance Specialisation (U Michigan)
    | M2 -- Risk and Return
    | Week 1 -- Bond Valuation
author: "Joe Nguyen"
date: "27 November 2015"
output: html_document
---

```{r, echo=FALSE}
rm(list = ls())

# Change working directory
dirBase <- "/home/joe/Documents/01-coursera/02-finance"
dirWorking <- "/02-risk-and-return/r-code"
setwd(paste0(dirBase, dirWorking))

# Generic finance functions
source(paste0(dirBase, "/01-tvm/r-code/func_finance.R"))
```

## 1.1 Financing: Bond and Stocks

| Assets | Liabilities |
| ------ | ----------- |
| **Real** assets | Equity / shares / stocks |
|                 | Debt / bonds |


## 1.2 Zero-Coupon Government Bonds

### Bonds

- A bond binds the borrower to the lender in a contract
- e.g. a loan
- Contract: entity (e.g. you) between entity (e.g. bank)


### Zero-Coupon (or Discount) Bonds

- This bond has only one payment, the face value, at maturity
- e.g. treasury bill or strip in the U.S.
- The price of a discount bond is `PV(P, r, n)`

#### Discount Bond: Example
A zero-coupon bond pays $1000 in exactly one year. What is the price of value (price of the bond) today if the interest rate on similar bonds is 2%?

```{r}
PV(1000, 0.02, 1)
```

#### Yield to Maturity: Example
What is the YTM of a 10-year zero-coupon bond with a face value of $1000 and a current price of $744.09?
```{r}
YTM(1000, 744.09, 10)

# or using IRR
cVec <- rep(0,10); cVec[10] <- 1000
IRR(744.09, cVec)
```

#### Yield to Maturity: Example 2
What is YTM if compounding interval is 6 months?
```{r}
YTM(1000, 744.09, 20)

# or using IRR
n <- 20
cVec <- rep(0,n); cVec[n] <- 1000
IRR(744.09, cVec)
```


## 1.3 Government Coupon Bonds

#### Coupon Bonds: Example
Suppose a government bond has a 6% coupon, a face value of $1000, and 10 years to maturity. What is the price of this bond given that similar bonds yield an annual return of (4, 6, 8)%?

Intepretation:
- Need to assume compounding interval is 6 months (gov. bond),
- "6% coupon" is a value (paid by gov. to you) every interval. The coupon price $C$ is:
$$
coupon \ percent = \frac{C}{fv}
$$

$$
6\% = \frac{C}{1000}
$$
So $C = \$60$ per year, which means $C = \$30$ per interval. Now, C is paid every period for 20 periods and the face value ($1000) is also paid at maturity.
- "What is price of this bond": How much is the PV of this bond, where you pay the face value ($1000) for it now?
```{r}
fv <- 1000
n <- 10 * halfsPerYear
r <- c(4,6,8,10,15,20) / 100 / halfsPerYear
c <- r[2] * fv
pvC <- annuity_pv(c, r, n)
pvFv <- PV(fv, r, n)
pvTot <- pvC + pvFv
rbind(pvC, pvFv, pvTot)

# Plot
nr <- length(r)
pvGroup <- factor(rep(1:3, each = nr), labels = c("pvC", "pvFv", "pvTot"))
pvDf <- data.frame(r = rep(r, times = 3), price = c(pvC, pvFv, pvTot), pvGroup)
library(ggplot2)
ggplot(data = pvDf, aes(r, price, group = pvGroup, col = pvGroup)) +
    geom_point(size = 4) + geom_line(lwd = 1.5)
```

So, if market value (interest rate $r$) is same as coupon rate, we buy bond at *face value* or at par. If $r < c_{rate}$, the bond is attractive to investors, who will purchase these bonds and bid the price up to a **premium** until $c_{rate} = r$. If interest rates are high $r > c_{rate}$, the bond is no longer attractive and the bond price will decrease, selling at a **discount** until $c_{rate} = r$.

Note however, that these prices adjust the bond face value according to $r$ and $n$ so that the bond is **fairly priced** according to TVM. There is no profit or loss.


## 1.5 Sources of Risk in Bonds

### Government Bonds

#### Interest rate risk -- two types:
1. Price risk (PR)
2. Coupon reinvestment risk (CRR)


### Corporate Bonds and Default Risk

- Corporate bonds almost always pay coupons
- Like government bonds, they are subject to interest rate risk
- But they are also subject to **default risk**






