---
title: |
    | Coursera Finance Specialisation (U Michigan)
    | M2 -- Risk and Return
    | Week 2 -- Stock Valuation
author: "Joe Nguyen"
date: "01 December 2015"
output: html_document
---

```{r, echo=FALSE}
rm(list = ls())

# Change working directory
dirBase <- "/home/joe/Documents/01-coursera/02-finance"
dirWorking <- "/02-risk-and-return/r-code"
setwd(paste0(dirBase, dirWorking))

# Generic finance functions
source(paste0(dirBase, "/01-tvm/r-code/func_finance.R"))
```


## 2.2 Stock Pricing / Valuation

- $P_0$ is today's stock price
- $P_1$ is the **expected** stock price next year
- Assume the expected dividend at the end of year is $DIV_1$

Now, if we sell the stock next year, it's price today is:
$$
P_0 = \frac{DIV_1 + P_1}{(1 + r)}
$$

Now what is $P_1$? It is simply $P_1 = \frac{DIV_2 + P_2}{(1 + r)}$, the time-adjusted value of the dividend and price at the end of year 2. By induction,
$$
\begin{align}
P_0 &= \frac{DIV_1 + \frac{DIV_2 + P_2}{(1 + r)}} {(1 + r)} \\
    &= \frac{DIV_1}{(1 + r)} + \frac{DIV_2}{(1 + r)^2} + \frac{P_2}{(1 + r)^2}
\end{align}
$$

Here, we assume the interest rate $r$ is constant across time. Now, in the limit where $n \rightarrow \infty$, the term $\frac{P_n}{(1 + r)^n} \rightarrow 0$. Thus, the stock price today is:
$$
\begin{align}
P_0 &= \sum_{t = 1}^n \frac{DIV_t}{(1 + r)^t} +  \frac{P_n}{(1 + r)^n} \\
    &= \sum_{t = 1}^{\infty} \frac{DIV_t}{(1 + r)^t}
\end{align}
$$


## 2.3 Dividend & Growth Stocks

### Special Case 1: Dividend Stock

Suppose dividends expect to remain approximately constant, what is the price?
$$
\begin{align}
P_0 &= \sum_{t = 1}^{\infty} \frac{DIV_t}{(1 + r)^t} \\
    &= DIV \sum_{t = 1}^{\infty} \frac{1}{(1 + r)^t} \\
\end{align}
$$

Let $S = \sum_{t = 1}^{\infty} \frac{1}{(1 + r)^t} = \frac{1}{(1 + r)} + \frac{1}{(1 + r)^2} + \ldots$. Then:
$$
\begin{align}
(1 + r) S &= 1 + \frac{1}{(1 + r)} + \frac{1}{(1 + r)^2} + \ldots \\
 S + rS   &= 1 + S \\
 S        &= \frac{1}{r}
\end{align}
$$

Thus,
$$
\boxed{ P_0 = \frac{DIV}{r} }
$$


#### Dividend Stock: Example

Suppose Green Utility is expected to pay a dividend of $0.50 per share for the foreseeable future and the return of this business is 10%. What should be the price of the stock?

```{r}
div <- 0.5
r <- 0.1
p0 <- div/r; p0
```
Interpret this as a **perpetuity**.

What if Green Utility paid the dividends for the next 30 years? What would the price be?
```{r}
n <- 30
p0 <- annuity_pv(div, r, n); p0
```
Interpret this as an **annuity**. Also, $0.29 is the PV at year 31 from year 31 to infinity.


### Special Case 2: Growth Stock

Suppose dividends are expected to grow at a rate of $g$ per year, what is the price of the stock?
$$
P_0 = \frac{DIV_1}{r - g}
$$

Derivation: using $DIV_2 = DIV_1 (1 + g)$ and so on,
$$
P_0 = \frac{DIV_1}{(1 + r)} + \frac{DIV_1 (1 + g)}{(1 + r)^2} + \ldots + \frac{DIV_1 (1 + g)^{n-1}}{(1 + r)^n} + \frac{P_n}{(1 + r)^n}
$$

In the limit, $\frac{P_n}{(1 + r)^n} = 0$. Define the infinite series $S$:
$$
S = \frac{DIV_1}{(1 + r)} + \frac{DIV_1 (1 + g)}{(1 + r)^2} + \ldots
$$

Now,
$$
\begin{align}
\frac{(1 + g)}{(1 + r)} S &= \frac{DIV_1 (1 + g)}{(1 + r)^2} + \ldots \\
\frac{(1 + g)}{(1 + r)} S &= S - \frac{DIV_1}{(1 + r)} \\
(1 + g) S                 &= (1 + r) S - DIV_1 \\
DIV_1                     &= S (1 + r - 1 - g) \\
S                         &= \frac{DIV_1}{r - g}
\end{align}
$$

#### Growth Stock: Example

Suppose Moggle is expected to pay dividends of $20 per share next year and the dividends are expected to grow indefinitely at a rate of 5% per year. Stocks of similar firms are earning an expected rate of return of 15% per year. What should be the price of a share of Moggle today?

```{r}
div <- 20
g <- 0.05
r <- 0.15
p0 <- div / (r - g); p0
```


## 2.4 The Engine of Growth (assume perpetuities)

- Invested capital per share (ICPS) - to start the enterprise
- Return on investment (ROI) > 0 for there to exist value creation / growth
- Cashflow per share: earnings per share (EPS) $EPS = ICPS \times ROI$. EPS comes from real assets, and is either paid out as dividends or reinvested:
    * Dividends (DIV)
    * Retained earnings (RE)
    
If we normalise by EPS:

* DIV / EPS = (1 - b)
* RE / EPS = b
    
then growth $g = b \times ROI$ when $b$ is reinvested into the enterpise / project.


## 2.5 Valuation / Growth: The Big Picture

THe price of a firm's stock can be divided up into:
$$
P_0 = \frac{EPS}{r} + PVGO
$$
where EPS is earnings (cashflow) per share and PVGO is the PV of growth opportunities. Now: $P_0 = PV(DIVs)$ with or without growth:

- $P_0 = \frac{DIV}{r}$
- $P_0 = \frac{DIV}{r - g}$


#### Good or bad growth

Suppose you know this abour Macrosoft, Inc.

- It is expected to earn 10% on its existing assets (ROI)
- It has $60 of capital per share
- The market capitalisation rate is 12% (r)
- It is not planning to grow

What should it's shares trade at?

```{r}
roi <- 0.1
eps <- 60 * roi
div <- eps
r <- .12

p0 <- div/r; p0
```


## 2.6 What is Growth?

#### New growth policy (for Macrosoft example): Scenario I

Suppose Macrosoft, Inc. announces an exciting growth policy that plows back 70% of its earnings every year at the same ROI as its existing assets.

- What growth rate will this policy generate?
- What will happen to its share price?

```{r}
# Growth rate
b <- 0.7
g <- b * roi; g * 100

# Share price
div <- (1 - b) * eps
p0 <- div / (r-g); p0
```


#### New growth policy (for Macrosoft example): Scenario II

Suppose instead Macrosoft, Inc. announces an alternative growth policy that plows back 50% of its earnings every year at an ROI of 14%.

- What growth rate will this policy generate?
- What will happen to its share price?

```{r}
# Growth rate
b <- 0.5
roi <- 0.14
g <- b * roi; g * 100

# Share price
#eps <- 60 * roi <= not applicable as new ROI on additional money plowed back only (according to discussion forum)
div <- (1 - b) * eps
p0 <- div / (r-g); p0
```



