---
title: "Week 4 Quiz"
author: "Joe Nguyen"
date: "08 December 2015"
output: html_document
---

```{r, echo=FALSE}
rm(list = ls())

# Change working directory
dirBase <- "/home/joe/Documents/01-coursera/02-finance"
dirWorking <- "/02-risk-and-return/r-code"
setwd(paste0(dirBase, dirWorking))

# Generic finance functions
source(paste0(dirBase, "/01-tvm/r-code/func_finance.R"))
```

## Question 1
(5 points) According to the principle of diversification, the only way to lower the risk of a portfolio you must add assets that are negatively correlated with your existing portfolio.

**FALSE** You can add stocks from other industries which may be positively or neutrally correlated because they may be correlated with the market movements. Additionally, you can diversify with defensive (non-cyclical) stocks.


## Question 2
(10 points) You have an equally weighted portfolio that consists of equity ownership in three firms. Firm A is trading at $62 share and has a beta of 0.82; Firm B is trading at $46 per share with a beta of 1.38; Firm C is trading at $134 per share with a beta of 0.64. Assume a risk free rate of 2.8% and market return of 7.0%. If each stock has a standard deviation of 20% and the stocks have a correlation of 0.50 with each other, your portfolio's expected return is closest to

```{r}
a.share <- 62
a.beta <- 0.82

b.share <- 46
b.beta <- 1.38

c.share <- 134
c.beta <- 0.64

rf <- 0.028
rm <- 0.07

sdAbc <- 0.2
corAbc <- 0.5
wAbc <- c(a.share, b.share, c.share); wAbc <- wAbc/sum(wAbc)

# Rate of return (on equity) for each stock
ri <- rf + (rm - rf) * c(a.beta, b.beta, c.beta)

# Portfolio r using equal weights
round(mean(ri) * 100, 2)

#Portfolio r using weights based on share price
round(ri %*% wAbc * 100, 2)

# # Portfolio risk (variance); weights are share prices
# pf.var <- sum(wAbc) * sdAbc^2 + 2*corAbc*(prod(wAbc[c(1,2)]) + prod(wAbc[c(1,3)]) + prod(wAbc[c(2,3)]))
# 
# pf.risk <- sqrt(pf.var)
# 
# # Rate of return (on equity) of portfolio
# rpf <- rf + (rm - rf) * pf.risk; rpf
```


## Question 3
(10 points) You have a portfolio that consists of equity ownership in three firms. You own 400 shares of Stout Drink Company (SDC), 500 shares of Carbon Computing (CC) and 400 shares of Serrano Foods (SF). Their current share prices are $80, $32, and $140, respectively. What is the percentage weight of SF in your portfolio? (No more than two decimals in the percentage but do not enter the % sign.)

```{r}
pf.n <- c(400, 500, 400)
pf.price <- c(80, 32, 140)

pfVal <- pf.n %*% pf.price

ast2w <- pf.n[2] * pf.price[2] / pfVal
round(ast2w * 100, 2)
```

### MARKED INCORRECT
A simple question, but an important one to sort out as an investor.

My mistake, should be [3] not [2]
```{r}
ast3w <- pf.n[3] * pf.price[3] / pfVal
round(ast3w * 100, 2)
```


## Question 4
(10 points) Historically, managed futures had a negative correlation with the S&P 500 index. In other words, managed futures tend to do poorly when the stock market outperforms. For stock investors seeking the portfolio diversification benefits, should they allocate any of their wealth to managed futures? 
**Yes** Diversification can include negatively correlated stocks.


## Question 5
(10 points) Suppose there are three securities (A, B, and C) to choose from to create your portfolio. Next year the economy will be in an expansion, normal, or recession state with probabilities 0.40, 0.33, and 0.27, respectively. The returns (%) on the securities in these states are as follows: Security A {expansion = +8, normal = +7, recession = +2}; Security B {-1, -1, +5}; Security C {+14, +7, -8}. You are considering 4 potential portfolios of these 3 securities, with the following specific weights on each: Portfolio I (0.24, 0.38, 0.38); Portfolio II (0.32, 0.34, 0.34); Portfolio III (0.48, 0.26, 0.26); Portfolio IV (0.72, 0.14, 0.14); where the numbers inside parentheses are weight of A, weight of B and weight of C, respectively. Which portfolio has the lowest risk?

```{r}
ps <- c(0.40, 0.33, 0.27)
ar <- c(8,7,2)
br <- c(-1,-1,5)
cr <- c(14,7,-8)
nStock <- 3

# Stocks returns
r <- matrix(c(ar,br,cr), nStock)

# Economy state factored stocks returns
rps <- r * ps

# Expected reward (of each stock)
reward <- colSums(rps); reward

# Expected risk (of each stock)
risk <- sqrt( colSums(ps * t((t(r) - reward)^2)) ); risk

# Covariance (ab, ac, bc)
combIdx <- matrix(c(1,2, 1,3, 2,3), 2)
covariance <- rep(0, nStock)
for (i in 1:nStock) {
    covariance[i] <- (rps[,combIdx[1,i]] - reward[combIdx[1,i]]) %*%
        (rps[,combIdx[2,i]] - reward[combIdx[2,i]])
}

# Portfolio weights
pfw1 <- c(0.24, 0.38, 0.38)
pfw2 <- c(0.32, 0.34, 0.34)
pfw3 <- c(0.48, 0.26, 0.26)
pfw4 <- c(0.72, 0.14, 0.14)
pfw <- matrix(c(pfw1,pfw2,pfw3,pfw4), 3)
npf <- 4

# Risk of each portfolio
pfVar <- colSums(pfw^2 * risk^2)
pfCovMat <- pfw[combIdx,]

covTerm <- rep(0, npf)
for (i in 1:nStock) {
    covTerm <- covTerm + 2 * apply(pfCovMat[(2*i-1):(2*i),], 2, prod) * covariance[i]
}
pfVar <- pfVar + covTerm

# Portfolio risk (for each portfolio)
sqrt(pfVar)
```
**Portfolio IV**


## Question 6
(10 points) The FTSE 100 is an index of the 100 largest market capitalization stocks traded on the London Stock Exchange. You think that 100 stocks are too much to keep up with, so you want to drop that number to 45. By doing this, what is the percentage drop in the UNIQUE relations between any two stocks in your portfolio that you will have to worry about? (No more than two decimals in the percentage but do not enter the % sign.)

```{r}
relUq100 <- choose(100,2)
relUq45 <- choose(45,2)
drop <- (relUq100 - relUq45) / relUq100
round(drop * 100, 2)
```


## Question 7
(10 points) Given the definition of beta in the context of the CAPM, a stock with low beta *can* have more volatility (variance or standard deviation) than a stock with high beta.

**FALSE** since beta is defined as the sensitivity to market risk, it measures volatility (i.e. risk). Thus, a high beta value equates to more volatility.

### MARKED INCORRECT
Think again, revisit the notion of systematic versus total risk of a stock.
Intuition for beta and total risk; risk of a stock in a large diversified portfolio versus risk of a stock by itself.

**From Investopedia:**

Beta is a measure of the volatility, or systematic risk, of a security or a portfolio in comparison to the market as a whole. Beta is used in the capital asset pricing model (CAPM), a model that calculates the expected return of an asset based on its beta and expected market returns.

Beta is calculated using regression analysis, and you can think of beta as the tendency of a security's returns to respond to swings in the market. A beta of 1 indicates that the security's price will move with the market. A beta of less than 1 means that the security will be less volatile than the market. A beta of greater than 1 indicates that the security's price will be more volatile than the market. For example, if a stock's beta is 1.2, it's theoretically 20% more volatile than the market.

Many utilities stocks have a beta of less than 1. Conversely, most high-tech, Nasdaq-based stocks have a beta of greater than 1, offering the possibility of a higher rate of return, but also posing more risk.

**Argument for answer TRUE**
Since beta captures historical/past risk of a stock, it is not useful for future predictions of the stock's movement (with respect to the market). Thus, a stock *could* previously have a low beta, but now be more volatile due to some event specific to the stock's industry (not market-wide).


## Question 8
(10 points) Suppose CAPM works, and you know that the expected returns on Google and IBM are estimated to be 13.50% and 10.75%, respectively. You have just calculated extremely reliable estimates of the betas of Google and IBM to be 1.22 and 0.80, respectively. Given this data, what is a reasonable estimate of the market risk-premium in percentage (the average/expected difference between the market return and the risk-free rate)? (No more than two decimals in the percentage but do not enter the % sign.)

```{r}
# return of c(google, IBM)
r <- c(0.135, 0.1075)
beta <- c(1.22, 0.8)

# least squares solution (exact in this case)
H <- cbind(beta, rep(1,2))
x <- solve(H, r)
# x <- lsfit(H, r)
rm <- x[1]
round(rm*100, 2)
```


## Question 9
(10 points) The standard deviation of a portfolio's return is the weighted average of the standard deviations of the returns of all securities in the portfolio, where the weights are proportional to the amount of your investment in a security relative to your total investment.

The standard deviation of a portfolio's return is its **risk**. The above statement is only true if all securities are pairwise correlated. Otherwise,
$$
\begin{align}
\sigma_p   &< \frac{1}{n} \sum \sigma_i \\
\sigma_p^2 &< \frac{1}{n^2} \left\{ \sum \sigma_i \right\}^2
\end{align}
$$

Look at **Week 3 Lecture Notes**. Answer is **FALSE**.


## Question 10
(15 points) Your company is very well-diversified, producing and selling all kinds of goods and services, with each division being approximately of the same value/size. The beta of your company is therefore effectively 1.00. The market risk premium (the average/expected difference between the market return and the risk-free rate) is 6.50% and the return on a long-term government bond is 4.25%. You have discovered an exciting opportunity to create a new medicine that will cure the common cold. This project will require a $1 billion investment, spread out equally between now (t = 0) and the end of this year (t = 1), and will produce $153 million dollars in perpetuity starting in year 2 (t = 2). Should you propose this investment to your CEO?

```{r}
beta <- 1
rf <- 0.0425
rPremium <- 0.065

# Project cost of capital
r <- rf + rPremium * beta

# NPV
capex <- 1e9
div <- 153e6
-capex + div/r
```

Since NPV is positive, **yes**, should propose this investment.

### MARKED INCORRECT
Rethink this issue before calculating stuff.
An issue that is more important to valuation than any other.

