---
title: "Final 1"
author: "Joe Nguyen"
date: "11 December 2015"
output: html_document
---

```{r, echo=FALSE}
rm(list = ls())

# Change working directory
dirBase <- "/home/joe/Documents/01-coursera/02-finance"
dirWorking <- "/02-risk-and-return/r-code"
setwd(paste0(dirBase, dirWorking))

# Generic finance functions
source(paste0(dirBase, "/01-tvm/r-code/func_finance.R"))
```

## Question 1
(10 points) Hard Spun Industries (HSI) has a project that it expects will produce a cash flow of $3.2 million in 14 years. To finance the project, the company needs to borrow $2.4 million today. The project will also produce intermediate cash flows of $240,000 per year that HSI can use to service coupon payments of $120,000 every six months. Based on the risk of this investment, market participants will require a 9.0% yield. If HSI wishes a maturity of 14 years (matching the arrival of the lump sum cash flow), what does the face value of the bond have to be? Recall that the compounding interval is 6 months and the YTM, like all interest rates, is reported on an annualized basis. (Enter just the number in dollars without the $ sign or a comma and round off decimals to the closest integer, i.e., rounding $30.49 down to $30 and rounding $30.50 up to $31.)

$$
price = \frac{c}{1 + r} + \frac{c}{(1 + r)^2} + \ldots + \frac{c}{(1 + r)^{28}} + \frac{faceVal}{(1 + r)^{28}}
$$
Rearrange for $faceVal$:
$$
faceVal = (1 + r)^{28} * \left( price - \text{annuity_pv(c, r, n)} \right)
$$

```{r}
price <- 2.4e6
cpay <- 120e3
r <- 0.09 / halfsPerYear
n <- 14 * halfsPerYear

faceVal <- (1 + r)^n * (price - annuity_pv(cpay, r, n))
round(faceVal)

faceVal <- FV(price - annuity_pv(cpay,r,n), r,n); round(faceVal)
```


## Question 2
(10 points) Five years ago, Highland, Inc. issued a corporate bond with an annual coupon of $5,000, paid at the rate of $2,500 every six months, and a maturity of 14 years. The par (face) value of the bond is $1,000,000. Recently, however, the company has run into some financial difficulty and has restructured its obligations. Today's coupon payment has already been paid, but the remaining coupon payments will be postponed until maturity. The postponed payments will accrue interest at an annual rate of 6.5% per year and will be paid as a lump sum amount at maturity along with the face value. The discount rate on the renegotiated bonds, now considered much riskier, has gone from 7.5% prior to the renegotiations to 14.0% per annum with the announcement of the restructuring. What is the price at which the new renegotiated bond should be selling today? Recall that the compounding interval is 6 months and the YTM, like all interest rates, is reported on an annualized basis. (Enter just the number in dollars without the $ sign or a comma and round off decimals to the closest integer, i.e., rounding $30.49 down to $30 and rounding $30.50 up to $31.)

```{r}
cpay <- 2500
n0 <- 14 * halfsPerYear
faceVal <- 1e6
r0 <- 0.075

g <- 0.065
r1 <- 0.14
n1 <- n0 - 5 * halfsPerYear

# Future value of remaining coupon payments
cpayFv <- annuity_fv(cpay, g, n1)
lump <- cpayFv + faceVal

# Price today
price <- PV(lump, r1, n1); round(price)
```

### Marked Incorrect
This is another example to make sure you understand and internalize the cash flow characteristics and pricing of bonds.

Forgot `/ halfsPerYear` for interest rates.

## Question 3
(10 points) Viento Windmills is a utility that charges customers for their wind generated electricity. With their current technology, they earn a total of $55 million each year to pay out to their 1 million shareholders. While their geographic footprint is fixed and we can't expect the climate to get progressively windier over time, they do have an opportunity to invest in technology that will more efficiently extract the wind energy and thus produce more megawatts to sell to customers. A one-time investment one year from now (t = 1) of $31 million for a state of the art lubricant system for the windmills will lead to extra cash flows that stay constant at $14 million per year starting the following year (t = 2) and lasting forever. If Viento plans to make the investment and cost of equity capital (the discount rate for equity) is 5.0%, calculate the increase or decrease in the share price of Viento as a result of this decision. Draw time line to understand what is going on. (Enter just the number in dollars without the $ sign or a comma and round off decimals to the closest integer, i.e., rounding $30.49 down to $30 and rounding $30.50 up to $31.)

```{r}
earn <- 55e6
nshares <- 1e6
eps <- earn / nshares

icost <- 31e6 / nshares

idiv <- 14e6 / nshares
r <- 0.05

# Nominal share price
p0 <- eps / r

## Share price with investment
# In year 1
div1 <- eps - icost

# From year 2 onwards
div2inf <- eps + idiv

# S -> infinite series starting with term 1 / (1 + r)^2 (dervied by hand)
S <- 1 / ( r^2 + r )
ip0 <- div1 / (1 + r) + div2inf * S

round(ip0 - p0)
```


## Question 4
(10 points) HigherEducation, Inc., a private educational company's share price is $130 per share; earnings and dividends are $6.60 a share, and the growth rate is zero. They have just announced a new growth strategy whereby the company's earnings would begin growing by 2% per year and remain stable at this new rate. This new growth strategy will require the company to reinvest 30% of their earnings starting at the end of this year (t = 1). What will happen to the price per share of this company? Think carefully, and draw a time line.

```{r}
p0 <- 130
div <- 6.60

g <- 0.02
b <- 0.3

# Original cost of capital
r <- div/p0

# Reduced dividend due to reinvestment
idiv <- div * (1 - b)

# New share price
ip <- idiv / (r - g)
round(ip - p0, 2)
```
**Price will increase by $20.15 per share.**


## Question 5
(15 points) You have just taken over as a fund manager at a brokerage firm. Your assistant, Thomas, is briefing you on the current portfolio and states "We have too much of our portfolio in Alpha. We should probably move some of those funds into Gamma so we can achieve better diversification." Is he right? Here is the data on all three stocks. Assume, for convenience, that all three securities do not pay dividends. Alpha, Current Price $30.00; Current Weight 90%; Next Year's Price: Expansion $36.00, Normal $32.50, Recession $25.00; Beta, Current Price 27.50; Current Weight 10%; Next Year's Price: Expansion 27.50, Normal 26, Recession 25; Gamma, Current Price $27.00; Current Weight 0%; Next Year's Price: Expansion $32.40, Normal $29.25, Recession $22.50. 

```{r}
a.p0 <- 30
a.w <- 0.9
a.p1 <- c(36, 32.5, 25)

b.p0 <- 27.5
b.w <- 0.1
b.p1 <- c(27.5, 26, 25)

c.p0 <- 27
c.w <- 0
c.p1 <- c(32.4, 29.25, 22.5)

# Determine (pairwise) correlation between a,c on p1
cor(a.p1, c.p1)
```
Since correlation is 1, **no** need to move funds from a to c.


## Question 6
(20 points) Suppose there are two mortgage bankers. Banker 1 has two $900,000 mortgages to sell. The borrowers live on opposite sides of the country and face an independent probability of default of 7%, with the banker able to salvage 46% of the mortgage value in case of default. Banker 2 also has two $900,000 mortgages to sell, but Banker 2's borrowers live on the same street, have the same job security and income. Put differently, the fates and thus solvency of Banker 2's borrowers move in lock step. They have a probability of defaulting of 3%, with the banker able to salvage 46% of the mortgage value in case of default. Both Bankers plan to sell their respective mortgages as a bundle in a mortgage-backed security (MBS) (i.e., as a portfolio). Which of the following is correct?

```{r}
# Banker 1 -- combined probability of default (due to independence)
0.07^2

# Banker 2 -- combined probability of default
0.03
```

Thus, Banker 1 has less risk than Banker 2. While both have equal return (46%) for each mortgage, Banker 1 has **lower expected return** because the probability of one default is higher (7% > 3%). Answer:

**Banker 1's MBS has a lower expected return and less risk.**

### Marked Incorrect

- Your rank of expected correct is valid but your rank of risk is wrong.
- A topical issue given the current crisis; requires you to both calculate and make decisions based on risk-return trade-offs.


## Question 7
(10 points) The beta of a portfolio is the weighted average of the betas of each security in the portfolio, where the weights are proportional to the amount of your investment in a security relative to your total investment.

Beta is risk (sd). If securities are not perfectly correlated, we know that portfolio risk is **less than** weighted average of the securities' risks.
$$
\sigma_p < \frac{1}{n} \sum \sigma_i
$$
Answer: **FALSE**

### Marked Incorrect

- Think carefully about CAPM and its form and determinants.
- Again, a basic but important, relation between the risks of portfolios and individual securities. Think about CAPM.


## Question 8
(15 points) Your company is very well-diversified, producing and selling all kinds of goods and services, with each division being approximately of the same value/size. The beta of your company is therefore effectively 1.00. The market risk premium (the average/expected difference between the market return and the risk-free rate) is 6.25% and the return on a long-term government bond is 3.75%. You have discovered an exciting opportunity to create a new medicine that will cure the common cold. This project will require a $1 billion investment, spread out equally between now (t = 0) and the end of this year (t = 1), and will produce $156 million dollars in perpetuity starting in year 2 (t = 2). Should you propose this investment to your CEO?

```{r}
price <- 1e9
cpay <- 156e6
rPrem <- 0.0625
rf <- 0.0375

r <- rf + rPrem * 1.0
p0 <- cpay / r

# Project return is positive (5.6e+08)
p0 - price
```

While the project return is positive ($560 million), **It can't be decided until the systematic risk of the project can be quantified.**


